Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-07-12 manuscript/manuscript.Rmd
adds section on SDMs to ms intro

@@

All R code and data necessary to reproduce our analysis is archived on GitHub as release 1.0 (http://github.com/atredennick/MicroMesoForecast/releases). That stable release will remain static as a record of this analysis, but subsequent versions may appear if we update this work.

@@

We fit all models using a hierarchical Bayesian approach, which we describe in more detail below. However, for each vital rate statistical model we also define the likelihood model we use. For the likelihood models, \textbf{Y} is always the relevant vector of observations (e.g., whether a  genet survived [1] or not [0] from year $t$ to $t+1$).

@@

\begin{align}
\text{logit}(S_{ijQ,t}) &= \gamma^{S}_{j,t} + \phi^{S}_{jQ} + \beta^{S}_{j,t}x_{ij,t} + \omega^{S}_{j}w_{ij,t} + \nu^{S}_{j}w_{ij,t}x_{ij,t} + \theta^{S}_{jk}C_{k,t} \\
Y^{S}_{ijQ,t} &\sim \text{Bernoulli}(S_{ijQ,t})
\end{align}

@@

\begin{align}
x_{ijQ,t+1} &= \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{ij,t} + \omega^{G}_{j}w_{ij,t} + \nu^{S}_{j}w_{ij,t}x_{ij,t} + \theta^{G}_{jk}C_{k,t} \\
Y^{G}_{ijQ,t} &\sim \text{Normal}(x_{ijQ,t+1}, \sigma_{j})
\end{align}

@@

\begin{equation}
Y^{R}_{jq,t+1} \sim \text{NegBin}(\lambda_{jq,t+1},\zeta)
\end{equation}

@@

\begin{align}
x_{jq,t+1} &= \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{jq,t} + \theta^{S}_{jk}C_{k,t} \\
Y^{G}_{jq,t+1} &\sim \text{LogNormal}(x_{jq,t+1}, \tau{j}) \text{T}[0,1]
\end{align}

@@

### Model fitting
Our Bayesian approach to fitting the vital rate models required choosing appropriate priors for unknown parameters and deciding which, if any, of those priors should be hierarchical. We decided to fit models where all terms were fit by species. Within a species, we fit yearly size effects and yearly intercepts hierarchically where year-specific coefficients were drawn from global distributions representing the mean size effect and intercept. We used flat, uninformative priors for all unknown parameters (Appendix X).

All of our analyses (model fitting and simulating) were conducted in R [@R2013]. We used the 'No-U-Turn' MCMC sampler in Stan [@stan2014]  to estimate the posterior distributions of model parameters using the package 'rstan' [@rstan2014]. We obtained posterior distributions for all model parameters from three parallel MCMC chains run for 1,000 iterations after discarding an initial 1,000 iterations. We recignize such short MCMC chains may surprise those more familiar with other MCMC samplers (i.e. JAGS or WinBUGS), but the Stan sampler is exceptionally efficient, which reduces the number of iterations needed to achieve convergence. We assessed convergence visually and made sure scale reduction factors for all parameters were less than 1.01. For the purposes of including parameter uncertainty in our population models, we saved the final 1,000 iterations from each of the three MCMC chains for all parameters to be used as randomly drawn values during population simulation. This step alleviates the need to reduce model parameters by model selection since sampling from the full parameter space in the MCMC ensures that if a parameter broadly overlaps zero, on average the effect in the population models will also be near zero.

@@

#### Quad-based model
Our quad-based model (QBM) perfectly mirrors its statistical description (Eq. 8-9). We use the same approach for drawing parameter values as described for the IPM.

@@

```{r figure_6, dependson="plot-options", fig.cap="Correlations between QBM and IPM estimates of climate effects. We ignore sizeXclimate interactions since these are not directly comparable across model types. The QBM does not have multiple vital rates, so its values are repeated across panels within each species.", fig.width=6, fig.height=6, cache=FALSE, results='hide'}
##  Figure script to look at correlations between IPM and QBM
##    climate effect estimates.

@@

ggplot(all_ests, aes(x=qbmvalue, y=value))+
  geom_point(size=3)+
  geom_smooth(method="lm", color="black", fill="grey")+
  facet_grid(species~vitalrate)+
  geom_text(data=cors, aes(label=paste("r = ", cor, sep="")), x=-1, y=1, size=4)+
  xlab("QBM Estimate")+
  ylab("IPM Estimate")+
  theme_bw()

>>>>

That population models do not scale easily explains, in part, why species distribution models (SDMs) have become a major tool for land managers and conservation planners. SDMs typically rely on easy-to-collect presence/absence data (but see...) that allow researchers to cover large spatial extents (e.g., ). Likewise, predictors in SDMs tend to be remotely sensed products. Thus, it is relatively straightforward to parameterize and project SDMs at landscape and regional scales, the very scales at which land managers work. However, the limitations of SDMs are well known (cite some review). SOME MORE STUFF... It seems that what land managers really want are population models at the landscape and regional scale, if only the ecological community could provide them.

@@

All R code and data necessary to reproduce our analysis is archived on GitHub as release v1.0[^revnote] (http://github.com/atredennick/MicroMesoForecast/releases). That stable release will remain static as a record of this analysis, but subsequent versions may appear if we update this work.

[^revnote]: *Note to reviewers*: so that v1.0 will be associated with the published version of the manuscript, we have released v0.1 to be associated with this review version.

@@

We fit all models using a hierarchical Bayesian approach. The models are fully descibed in Appendix A, so here we focus on the main process and the mode likelihood. For the likelihood models, $\textbf{y}^X$ is always the relevant vector of observations for vital rate *X* ($X = S, G, R, or P for survival, growth, recruitment, or population growth) For example, $\textbf{y}^S$ is a vector of 0s and 1s indicating whether a genet surives from *t* to *t+1*, or not.

@@

\begin{align}
\text{logit}(S_{ijQ,t}) &= \gamma^{S}_{j,t} + \phi^{S}_{jQ} + \beta^{S}_{j,t}x_{ij,t} + \omega^{S}_{j}w_{ij,t} + \nu^{S}_{j}w_{ij,t}x_{ij,t} + \theta^{S}_{jk}C_{k,t} \\
y^{S}_{ijQ,t} &\sim \text{Bernoulli}(S_{ijQ,t})
\end{align}

@@

\begin{align}
x_{ijQ,t+1} &= \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{ij,t} + \omega^{G}_{j}w_{ij,t} + \nu^{S}_{j}w_{ij,t}x_{ij,t} + \theta^{G}_{jk}C_{k,t} \\
y^{G}_{ijQ,t} &\sim \text{Normal}(x_{ijQ,t+1}, \sigma_{j})
\end{align}

@@

\begin{equation}
y^{R}_{jq,t+1} \sim \text{NegBin}(\lambda_{jq,t+1},\zeta)
\end{equation}

@@

\begin{align}
x_{jq,t+1} &= \gamma^{P}_{j,t} + \phi^{P}_{jQ} + \beta^{P}_{j,t}x_{jq,t} + \theta^{P}_{jk}C_{k,t} \\
y^{P}_{jq,t+1} &\sim \text{LogNormal}(x_{jq,t+1}, \tau{j}) \text{T}[0,1]
\end{align}

@@

### Model fitting
Our Bayesian approach to fitting the vital rate models required choosing appropriate priors for unknown parameters and deciding which, if any, of those priors should be hierarchical. We decided to fit models where all terms were fit by species. Within a species, we fit yearly size effects and yearly intercepts hierarchically where year-specific coefficients were drawn from global distributions representing the mean size effect and intercept. We used uninformative priors (Appendix A).

All of our analyses (model fitting and simulating) were conducted in R [@R2013]. We used the 'No-U-Turn' MCMC sampler in Stan [@stan2014]  to estimate the posterior distributions of model parameters using the package 'rstan' [@rstan2014]. We obtained posterior distributions for all model parameters from three parallel MCMC chains run for 1,000 iterations after discarding an initial 1,000 iterations. We recignize such short MCMC chains may surprise those more familiar with other MCMC samplers (i.e. JAGS or WinBUGS), but the Stan sampler is exceptionally efficient, which reduces the number of iterations needed to achieve convergence. We assessed convergence visually and made sure scale reduction factors for all parameters were less than 1.01. For the purposes of including parameter uncertainty in our population models, we saved the final 1,000 iterations from each of the three MCMC chains for all parameters to be used as randomly drawn values during population simulation. This step alleviates the need to reduce model parameters by model selection since sampling from the full parameter space in the MCMC ensures that if a parameter broadly overlaps zero, on average the effect in the population models will also be near zero. We report the posterior mean, standard deviation, and 95% Bayesian Credible Intervals for every parameter of each model for each species in Appendix B.

@@

#### Quad-based model
Our quad-based model (QBM) perfectly mirrors its statistical description (Eqs. 8-9). We use the same approach for drawing parameter values as described for the IPM.

@@

```{r figure_6, dependson="plot-options", fig.cap="Correlations between QBM and IPM estimates of climate effects. We ignore sizeXclimate interactions since these are not directly comparable across model types. The QBM does not have multiple vital rates, so its values are repeated across panels within each species. Across top panels, 'growth' = growth regression, 'rec' = recruitment regression, 'surv' = survival regression.", fig.width=6, fig.height=6, cache=FALSE, results='hide'}
##  Figure script to look at correlations between IPM and QBM
##    climate effect estimates.

@@

ggplot(all_ests, aes(x=qbmvalue, y=value))+
  geom_point(size=3)+
  geom_smooth(method="lm", color="black", fill="grey", se=FALSE)+
  facet_grid(species~vitalrate)+
  geom_text(data=cors, aes(label=paste("r = ", cor, sep="")), x=-1, y=1, size=4)+
  xlab("QBM Estimate")+
  ylab("IPM Estimate")+
  theme_bw()

### 2015-07-12 manuscript/manuscript.tex
adds section on SDMs to ms intro

@@

All R code and data necessary to reproduce our analysis is archived on
GitHub as release 1.0
(\url{http://github.com/atredennick/MicroMesoForecast/releases}). That
stable release will remain static as a record of this analysis, but
subsequent versions may appear if we update this work.

@@

We fit all models using a hierarchical Bayesian approach, which we
describe in more detail below. However, for each vital rate statistical
model we also define the likelihood model we use. For the likelihood
models, \textbf{Y} is always the relevant vector of observations (e.g.,
whether a genet survived {[}1{]} or not {[}0{]} from year $t$ to $t+1$).

@@

Our Bayesian approach to fitting the vital rate models required choosing
appropriate priors for unknown parameters and deciding which, if any, of
those priors should be hierarchical. We decided to fit models where all
terms were fit by species. Within a species, we fit yearly size effects
and yearly intercepts hierarchically where year-specific coefficients
were drawn from global distributions representing the mean size effect
and intercept. We used flat, uninformative priors for all unknown
parameters (Appendix X).

All of our analyses (model fitting and simulating) were conducted in R
(R Core Development Team 2013). We used the `No-U-Turn' MCMC sampler in
Stan (Stan Development Team 2014a) to estimate the posterior
distributions of model parameters using the package `rstan' (Stan
Development Team 2014b). We obtained posterior distributions for all
model parameters from three parallel MCMC chains run for 1,000
iterations after discarding an initial 1,000 iterations. We recignize
such short MCMC chains may surprise those more familiar with other MCMC
samplers (i.e.~JAGS or WinBUGS), but the Stan sampler is exceptionally
efficient, which reduces the number of iterations needed to achieve
convergence. We assessed convergence visually and made sure scale
reduction factors for all parameters were less than 1.01. For the
purposes of including parameter uncertainty in our population models, we
saved the final 1,000 iterations from each of the three MCMC chains for
all parameters to be used as randomly drawn values during population
simulation. This step alleviates the need to reduce model parameters by
model selection since sampling from the full parameter space in the MCMC
ensures that if a parameter broadly overlaps zero, on average the effect
in the population models will also be near zero.

@@

Our quad-based model (QBM) perfectly mirrors its statistical description
(Eq. 8-9). We use the same approach for drawing parameter values as
described for the IPM.

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_6.pdf}
\caption{Correlations between QBM and IPM estimates of climate effects.
We ignore sizeXclimate interactions since these are not directly
comparable across model types. The QBM does not have multiple vital
rates, so its values are repeated across panels within each species.}
\end{figure}

>>>>

That population models do not scale easily explains, in part, why
species distribution models (SDMs) have become a major tool for land
managers and conservation planners. SDMs typically rely on
easy-to-collect presence/absence data (but see\ldots{}) that allow
researchers to cover large spatial extents (e.g., ). Likewise,
predictors in SDMs tend to be remotely sensed products. Thus, it is
relatively straightforward to parameterize and project SDMs at landscape
and regional scales, the very scales at which land managers work.
However, the limitations of SDMs are well known (cite some review). SOME
MORE STUFF\ldots{} It seems that what land managers really want are
population models at the landscape and regional scale, if only the
ecological community could provide them.

@@

All R code and data necessary to reproduce our analysis is archived on
GitHub as release v1.0\footnote{\emph{Note to reviewers}: so that v1.0
  will be associated with the published version of the manuscript, we
  have released v0.1 to be associated with this review version.}
(\url{http://github.com/atredennick/MicroMesoForecast/releases}). That
stable release will remain static as a record of this analysis, but
subsequent versions may appear if we update this work.

@@

We fit all models using a hierarchical Bayesian approach. The models are
fully descibed in Appendix A, so here we focus on the main process and
the mode likelihood. For the likelihood models, $\textbf{y}^X$ is always
the relevant vector of observations for vital rate \emph{X}
($X = S, G, R, or P for survival, growth, recruitment, or population growth) For example, $\textbf{y}\^{}S\$
is a vector of 0s and 1s indicating whether a genet surives from
\emph{t} to \emph{t+1}, or not.

@@

Our Bayesian approach to fitting the vital rate models required choosing
appropriate priors for unknown parameters and deciding which, if any, of
those priors should be hierarchical. We decided to fit models where all
terms were fit by species. Within a species, we fit yearly size effects
and yearly intercepts hierarchically where year-specific coefficients
were drawn from global distributions representing the mean size effect
and intercept. We used uninformative priors (Appendix A).

All of our analyses (model fitting and simulating) were conducted in R
(R Core Development Team 2013). We used the `No-U-Turn' MCMC sampler in
Stan (Stan Development Team 2014a) to estimate the posterior
distributions of model parameters using the package `rstan' (Stan
Development Team 2014b). We obtained posterior distributions for all
model parameters from three parallel MCMC chains run for 1,000
iterations after discarding an initial 1,000 iterations. We recignize
such short MCMC chains may surprise those more familiar with other MCMC
samplers (i.e.~JAGS or WinBUGS), but the Stan sampler is exceptionally
efficient, which reduces the number of iterations needed to achieve
convergence. We assessed convergence visually and made sure scale
reduction factors for all parameters were less than 1.01. For the
purposes of including parameter uncertainty in our population models, we
saved the final 1,000 iterations from each of the three MCMC chains for
all parameters to be used as randomly drawn values during population
simulation. This step alleviates the need to reduce model parameters by
model selection since sampling from the full parameter space in the MCMC
ensures that if a parameter broadly overlaps zero, on average the effect
in the population models will also be near zero. We report the posterior
mean, standard deviation, and 95\% Bayesian Credible Intervals for every
parameter of each model for each species in Appendix B.

@@

Our quad-based model (QBM) perfectly mirrors its statistical description
(Eqs. 8-9). We use the same approach for drawing parameter values as
described for the IPM.

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_6.pdf}
\caption{Correlations between QBM and IPM estimates of climate effects.
We ignore sizeXclimate interactions since these are not directly
comparable across model types. The QBM does not have multiple vital
rates, so its values are repeated across panels within each species.
Across top panels, `growth' = growth regression, `rec' = recruitment
regression, `surv' = survival regression.}
\end{figure}
